---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0.1
#ESSENTIALS
title: 'Use of management strategy evaluation to inform in fisheries management in a changing climate'
author:
  - name: '**Jonathan W. Cummings**'
    affil: 1
    main: true
    orcid: '0000-0001-8028-5787'
    twitter: jonwcummings 
    email: 'jcummings@umassd.edu'
  - name: Amanda Hart
    affil: 1
    main: true
    orcid: '0000-0002-6449-9328'
    email: 'ahart2@umassd.edu'
  - name: Gavin Fay
    main: true
    affil: 1
    email: 'gfay@umassd.edu'
affiliation:
  - num: 1
    address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
main_findings:
  - "Make **better posters** with RMarkdown + **posterdown**."
logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logoright_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logocenter_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/qr-code-black.png
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: packages.bib
link-citations: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

# Introduction

For most of its history fisheries management has taken place against the backdrop of ecosystem status with fishing impacts occupying center stage.  However, the increased rate and degree of ecosystem climatic change is also bringing ecosystem status to the forefront of fisheries management.
Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” (citation). Thus, MSE is a compelling tool to address the impact of climate change on fisheries management.
Despite numerous applications of MSE, the application that include climate change are limited, leaving uncertainty about the best approach to applying MSEs that account for a changing environment.


```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

## Objectives

Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives. With an eye towards adaptive management, we seek to note the lessons learned from successful MSEs to:

* Highlight unique aspects of MSEs that model climate change as a driver
* Identify shortcomings in MSE documentation that delays learning
* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate

Need climate change imagery

Need MSE imagery

Need learning and adaptive management imagery

# Methods

I will show here how to include poster elements that may be useful, such as an equation using mathjax:

$$
E = mc^2
$$

To reference a citation you can add your `.bib` file to the working directory and name it in the YAML metadata or generate an automated one as done here, then you only need to reference the label value in the `.bib` file. For example this package is built on top of the wonderful {pagedown} package and I will cite it at the end of this sentance using this in the rmd `[@R-pagedown]` [@R-pagedown].

To get a better understanding of how to include features like these please refer to the {posterdown} [wiki](https://github.com/posterdown/wiki).

**_Now on to the results!_**

# Results

Here you may have some figures to show off, bellow I have made a scatterplot with the infamous Iris dataset and I can even reference to the figure automatically like this, `Figure \@ref(fig:irisfigure)`, Figure \@ref(fig:irisfigure).

```{r, irisfigure, fig.cap='Here is a caption for the figure. This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="80%"}
par(mar=c(2,2,0,1))
plot(x = iris$Sepal.Length, y = iris$Sepal.Width,
     col = iris$Species, pch = 19, xlab = "Sepal Length",
     ylab = "Sepal Width")
```

Maybe you want to show off some of that fancy code you spent so much time on to make that figure, well you can do that too! Just use the `echo=TRUE` option in the r code chunk options, Figure \@ref(fig:myprettycode)!

```{r myprettycode, echo=TRUE, fig.cap='Boxplots, so hot right now!', fig.height=6, out.width="80%"}
#trim whitespace
par(mar=c(2,2,0,0))
#plot boxplots
boxplot(iris$Sepal.Width~iris$Species,
        col = "#008080", 
        border = "#0b4545",
        ylab = "Sepal Width (cm)",
        xlab = "Species")
```

How about a neat table of data? See, Table \@ref(tab:iristable):

```{r, iristable}
knitr::kable(
  iris[1:15,1:5], format = "html",
  caption = "A table made with the **knitr::kable** function.",
  align = "c", col.names = c("Sepal <br> Length",
                             "Sepal <br> Width",
                             "Petal <br> Length",
                             "Petal <br> Width",
                             "Species"),
  escape = FALSE)
```

# References
